---
title: "Lab 03 - Nobel laureates"
author: "Enhui Wang"
date: "2025.1.23"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1
Nobel has 935 observations and 26 variables.Each row presents a Nobel laureate such as names, the year they won,biographical information, and etc.

```{r}
nrow(nobel)#the number of rows (observations)
ncol(nobel)#the number of columns (variables)
```

### Exercise 2

```{r}
nobel_living <- nobel %>%
  filter(!is.na(country),#laureates for whom country is available
  gender != "org", #exclude organizations
  is.na (died_date))# still live people

#check the number of observations
nrow(nobel_living)
```

### Exercise 3
According to the plot, the Buzzfeed headline is supported by the data because the US bars are significantly taller across all categories.

```{r}
#Identify whether the laureate was in the USA
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  ) 

#limit the analysis to some categories
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r bar_plot}
library(ggplot2)

#create the faceted bar plot
ggplot(nobel_living_science, aes(x = country_us, fill = country_us )) + 
  geom_bar() +
  facet_wrap (~ category) + #creates separate facets for each category
  coord_flip() + #flips the bars horizontally
  labs(
  title = "Nobel Laureates by Prize Category and Location",
  x = "Location",
  y = "Number",
  fill = "Location" )
```





### Exercise 4
105 winners are born in the US.
```{r}
#Create a new variable called born_country_us
nobel_living_science <- nobel_living_science %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  ) 

#Count the number of winners born in the US
us_born_number <- nobel_living_science %>%
  filter(born_country_us == "USA") %>%
  summarise(count = n())

us_born_number
```

### Exercise 5
It supports Buzzfeed's claim because the "USA" bars are longer than "Others" in most facets, excluding Chemistry
ore
```{r}
library(ggplot2)

#create the faceted bar plot
ggplot(nobel_living_science, aes(x = country_us, fill = born_country_us )) + 
  geom_bar(position = "stack") +
  facet_wrap (~ category) + #creates separate facets for each category
  coord_flip() + #flips the bars horizontally
  labs(
  title = "Nobel Laureates by Prize Category, win_prize_place, and birthplace",
  x = "Location",
  y = "Number",
  fill = "Born in" )
```

### Exercise 6

...
