---
title: "Lab 03 - Nobel laureates"
author: "Enhui Wang"
date: "2025.1.23"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1
Nobel has 935 observations and 26 variables.Each row presents a Nobel laureate such as names, the year they won,biographical information, and etc.

```{r}
nrow(nobel)#the number of rows (observations)
ncol(nobel)#the number of columns (variables)
```

### Exercise 2

```{r}
nobel_living <- nobel %>%
  filter(!is.na(country),#laureates for whom country is available
  gender != "org", #exclude organizations
  is.na (died_date))# still live people

#check the number of observations
nrow(nobel_living)
```

### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
